\section{Line-search methods}
\begin{slide}[Generic Line Search]
Given an uncontrained minimization problem $$x^* = \operatorname{arg\ min}_x f(x)$$, the methods of line search attempt to find the optimum point through the following algorithm.
\begin{description}
	\item[Step 1] Find an good descend direction $d_k$
	\item[Step 2] Solve $\min_\alpha f(x_k +\alpha d_k)$
	\item Repeat until convergence.
\end{description}
Line search methods are distinguished  based on how step 2 are implemented: Exact Line search and Inexact line search
\end{slide}

\begin{slide}[Exact Line Search]
\begin{list}{}{\setlength\labelwidth{0cm}}
\item \textbf{Line Search by interval shrinking}
	Given an interval $[a,b]$, find which interval has minimum $[a,a+\tau(b-a)]$ or $[b-\tau(b-a),b]$
	\item Bisection method: $\tau = 1/2$
	\item Fibonacci search: $\tau = {F_{k-1}/F_k}$.
	\item Golden Section: $\tau = {1-\sqrt{5} \over 2}$
\item \textbf{Line search by curve fitting}
	Fit a curve to the function and then find the minimum point of the curve.
	\item Newtons method: Fit a quadratic curve. Need $f(x_k)$, $f'(x_k)$, and $f''(x_k)$
	\item Method of False position: Use $f'(x_{k-1}) - f'(x_k) \approx f''(x_k)$
	\item Cubic Fit: Fit a cubic curve through  $f(x_k), f(x_{k-1}),f'(x_k), f'(x_{k-1})$
	\item Quadratic Fit: fit a quadratic function through $f(x_1),f(x_2), f(x_3)$
\end{list}
\end{slide}
\begin{slide}[Inexact Line-search]
	TEst which ensure an selected step size is ok!

	Define $\phi(\alpha) = f(x_k +\alpha d_k)$

	\textbf{Armijio rule}
		$$\phi(\alpha) < \phi(0) + \epsilon \phi'(0)\alpha \qquad 0<\epsilon<1$$
		Selected step $\alpha$ size is good if the function value after taking that step is below the line in equation. Practically, $\epsilon = 0.2$
		\footnotetext{$\phi(0) + \phi'(0)\alpha$ is first order approximation of $\phi(\alpha)$ around $\alpha = 0$}
	
	\textbf{Wolfe Test}
		$$\phi'(\alpha) \geq (1-\epsilon)\phi'(0)$$
			There is a change in curvature. Note that Wolfe test requires the step size to satisfy Armijio rule first but $0<\epsilon<0.5.$

	\textit{Backtracking Line Search}: $\alpha = 1$, $0<\epsilon<1$ and $\eta>1$. Do $\alpha = \eta \alpha$ until Armijio rule is satisfied.
\end{slide}
